---
output:
  md_document:
    variant: markdown_github
  bibliography: ./bibliography.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# hyetor

  [![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/kvantas/hyetor?branch=master&svg=true)](https://ci.appveyor.com/project/kvantas/hyetor)
  [![Travis build status](https://travis-ci.org/kvantas/hyetor.svg?branch=master)](https://travis-ci.org/kvantas/hyetor)
  [![Coverage status](https://codecov.io/gh/kvantas/hyetor/branch/master/graph/badge.svg)](https://codecov.io/github/kvantas/hyetor?branch=master)
  [![DOI](https://zenodo.org/badge/145962234.svg)](https://zenodo.org/badge/latestdoi/145962234)
  [![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

`hyetor` is an R  package that provides a collection of tools that analyze fixed interval precipitation records. It can be used to:

 1. Aggregate precipitation values.
 2. Split precipitation time series to independent rainstorms using the maximum dry period duration of no precipitation.
 3. Compile Unitless Cumulative Hyetographs.
 4. Find maximum rainfall intensities.
 5. Categorize rainstorms using Huff's quartile classification.
 6. Calculate rainfall erosivity values
 7. Create missing values summaries.

For more details checkout the package's [ website](https://kvantas.github.io/hyetor/) and the
vignettes:

 
## Installation

You can install the development version from Github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("kvantas/hyetor")
```

## Using hyetor

The functions that are provided by `hyetor` are:

* Functions that can be used to preprocess precipitation time-series: `hyet_create`, `hyet_fill`, `hyet_aggregate` and `hyet_split`.
* Functions that analyze precipitation time-series: `hyet_erosivity`, `hyet_intensities`, `hyet_missing` and `hyet_uch`.

The data sets that are provided by `hyetor` are:

* `prec5min`, This time series comes from the weather station 'Arna' in Greece. The owner of that weather station is the Ministry of Environment and Energy. The time series period is from 12/1954 to 05/1956 and the time-step is 5 minutes.

## Example

This is a minimal example which shows how to use the package's functions to analyze the internal data set and compute erosivity values.

Load libraries and view data:
```{r, "load libraries", message=FALSE}
library(hyetor)
library(tibble)
library(dplyr)
library(lubridate)

prec5min
```

Compute the missing values ratio per year:
```{r "missing values report"}
prec5min %>%
  hyet_fill(time_step = 5, ts_unit = "mins") %>%
  hyet_missing() %>%
  group_by(year) %>%
  summarise(na_ratio = mean(na_ratio))
```

Split to independent rainstorms and calculate erosivity values with cumulative precipitation greater than 12,7 mm:

```{r "compute erosivity"}
prec5min %>%
  hyet_split(time_step = 5, ts_unit = "mins") %>%
  hyet_erosivity(time_step = 5) %>%
  ungroup() %>%
  filter(cum_prec > 12.7) %>%
  select(begin, duration, cum_prec, erosivity)
  
```



## Meta

* Bug reports, suggestions, and code are welcome. Please see [Contributing](/CONTRIBUTING.md).
* Licence:
    + All code is licensed MIT.
    + All data are from the public data sources in http://www.hydroscope.gr/.
* To cite `hyetor` please use:

```
  Vantas, (2018). hyetor: R package to analyze fixed interval precipitation time series, URL: https://kvantas.github.io/hyetor/,
  DOI:http://doi.org/10.5281/zenodo.1403156
```
A BibTeX entry for LaTeX users is
```
@Manual{ vantas2018hyetor,
    author = {Konstantinos Vantas},
    title = {{hyetor}: R package  to analyze fixed interval precipitation time series},
    doi = {http://doi.org/10.5281/zenodo.1403156},
    year = {2018},
    note = {R package},
    url = {https://kvantas.github.io/hyetor/},
  }
```
