---
title: "Rainfall erosivity calculation"
author: "Konstantinos Vantas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction

The R coefficient (MJ.mm/ha/h/yr) is defined as the long-term average of the product of the kinetic energy of a storm and the maximum 30 min intensity [@renard1991rusle]:

$$R = \frac{1}{n} \sum_{j=1}^{n} \sum_{k=1}^{m_j} (EI_{30})_{k}$$

where

 * $n$ is the number of years with rainfall records,
 * $m_j$ is the number of storms during year $j$ and
 * $EI_{30}$ is the erosivity of storm $k$. 
 
The erosivity $EI_{30}$ (MJ.mm/ha/h) is equal to:

$$EI_{30} = \left( \sum_{r=1}^{m} e_r \cdot v_{r} \right) \cdot I_{30}$$

where:

* $e_r$ is the energy of rainfall (MJ/ha/mm),
* $v_r$ is the rainfall depth (mm) for the time interval $r$ of the hyetograph, which has been divided into $r = 1, 2, ..., m$ sub-intervals, such that each one of these is characterized by constant rainfall intensity and
* $I_{30}$ is the maximum rainfall intensity for a 30 minutes duration. 

The quantity $e_r$ can be calculated for each $r$ using one of the kinetic energy equations:

1. Wischmeier and Smith equation, used in USLE: $e_r = 0.119 + 0.0873log_{10}(i)$ with the upper limit of 0.283  MJ/ha/mm if ${i} > 76$ mm/h. [@wischmeier1958rainfall].
2. Brown and Foster equation, used in RUSLE $e_r = 0.29(1 - 0.72  e^{-0.05i})$ [@brown1987storm].
3. McGregor et al. equation used in RUSLE2 $e_r = 0.29(1 - 0.72  e^{-0.082i})$ [@McGregor1995].

In the above equations $i_r$ is the rainfall intensity (mm/hr) and $e_r$ is the kinetic energy per unit of rainfall (MJ/ha/mm) for the interval $r$.

The rules that apply in order to single out the storms causing erosion and to divide rainfalls of large duration are:

1. A rainfall event is divided into two parts, if its cumulative depth for duration of 6 hours at a certain location is less than 1.27 mm.
2. A rainfall is considered erosive:
   * if it has a cumulative value greater than 12.7 mm or
   * during a time period of 15 mins a cumulative value of precipitation of at least 6.4 mm is recorded.

## Example

This is an example that uses the internal data set in order to compute the corresponding rainfall erosivity values.

```{r, "load libraries", message=FALSE}
library(hyetor)
library(tibble)
library(dplyr)
library(lubridate)

# view data
prec5min
```

The following code can be used to: 

a) Fill the time-series. 
b) Compute the rainfall erosivity values per storm.
c) Filter the above values using cumulative precipitation height and maximum 15 minutes intensity rules.

```{r "compute erosivity"}
ei_values <- prec5min %>%
  hyet_fill(time_step = 5, ts_unit = "mins") %>%
  hyet_erosivity(time_step = 5) %>%
  filter(cum_prec > 12.7 | max_i15 > 4 * 6.4)

ei_values
  
```

After the calculation of $EI30$ values the $R$ coefficient can be computed with:

```{r compute R values}

# add years and months variables
ei_values <- ei_values %>%
  mutate( year = year(begin))

# compute R
R_coeff <- ei_values %>%
  group_by(year) %>%
  summarise(R = sum(erosivity)) %>%
  ungroup() %>%
  summarise(R = mean(R)) %>%
  unlist()

R_coeff

```


## References


